\vspace{-2.0cm}The effect of nitrogen application time on infection and development of the Wheat Streak Mosaic Virus (WSMV)

### 2015-09-30 draft1leslieandkenny.Rnw
Some code tweaks. (Simulated data using NBin.)

# Combine nitrogen and inoc/cntl
fert2$n.status <- with(fert2, interaction(n.trt, status))
@

<<fakedata,echo=FALSE>>=
# Create some fake data with no differences anywhere
set.seed(983724)
fert.sim1 <- data.frame(
  'block' = c(rep('A', 30), rep('B', 30), rep('C', 30), rep('D', 30), rep('E', 30), rep('F', 30)),
  'variety' = c(rep('Var1', 6), rep('Var2', 6), rep('Var3', 6), rep('Var4', 6), rep('Var5', 6)),
  'n.application' = rep(c('FALL', 'FALL', 'EARLY SPRING', 'EARLY SPRING', 'LATE SPRING', 'LATE SPRING'), 30),
  'status' = rep(c('CNTL', 'INOC'), 90),
  'total' = rep(30, 180),
  'infected' = sample(fert2$infected, 180, replace = TRUE))

@@

# Create a row factor
fert.sim1$row <- with(fert.sim1, interaction(variety, block))
@

@@

The model is
$$
y_i\sim\mathrm{Binomial}(30, p_i),
$$
$$
\mathrm{logit}(p_i)=\mathbf{x}_i\mathbf{\beta}+b_{0,j[i]}+b_{1,j[i],k[i]}
$$
where
\begin{itemize}
\item \(p_i\) is the probability of infection for the \(i\)th observation,
\item \(\mathrm{logit}(p_i)=\log\left(\dfrac{p_i}{1-p_i}\right)\) is the
natural logarithm of the odds of infection,
\item \(\mathbf{x_i}\) is a row vector containing indicator variables for
the variety, nitrogen application timing, and inoculation status of
observation \(i\), and
\item \(b_{0,j}\sim\mathrm{N}(0,\sigma^2_{b_0})\) and
\(b_{1,j,k}\sim\mathrm{N}(0,\sigma^2_{b_1})\) are random effects for block
and row, respectively.
\end{itemize}

@@

<<models,cache=TRUE>>=
glmm3way <- glmer(cbind(infected, total) ~ variety*n.trt*status +
                    (1|block/row),
                  control = glmerControl(optimizer = 'bobyqa'),
                  family = binomial, data = fert.sim1)
summary(glmm3way)

glmm2way <- glmer(cbind(infected, total) ~ variety*n.trt + variety*status +
                    n.trt*status + (1|block/row),
                    control = glmerControl(optimizer = 'bobyqa'),
                  family = binomial, data = fert.sim1)
summary(glmm2way)

>>>>

# Combine nitrogen and inoc/cntl (for the heatmap labels)
fert2$n.status <- with(fert2, interaction(n.trt, status))
@

<<fakedata,echo=FALSE>>=
# Create some fake data with no differences anywhere
set.seed(983724)
fert.sim1 <- data.frame(
  'block' = c(rep('A', 30), rep('B', 30), rep('C', 30), rep('D', 30), rep('E', 30), rep('F', 30)),
  'variety' = c(rep('Var1', 6), rep('Var2', 6), rep('Var3', 6), rep('Var4', 6), rep('Var5', 6)),
  'n.application' = rep(c('FALL', 'FALL', 'EARLY SPRING', 'EARLY SPRING', 'LATE SPRING', 'LATE SPRING'), 30),
  'status' = rep(c('CNTL', 'INOC'), 90),
  'total' = rep(30, 180),
  'infected' = rnbinom(180, 4, 7/9))
#  'infected' = sample(fert2$infected, 180, replace = TRUE))

@@

# Create a row factor
fert.sim1$row <- with(fert.sim1, interaction(variety, block))

# Combine nitrogen and inoc/cntl
fert.sim1$n.status <- with(fert.sim1, interaction(n.trt, status))
@

@@

The model is
$$
y_i\sim\mathrm{Binomial}(30, p_i),
$$
$$
\mathrm{logit}(p_i)=\mathbf{x}_i\mathbf{\beta}+b_{0,j[i]}+b_{1,j[i],k[i]}
$$
where
\begin{itemize}
\item \(p_i\) is the probability of infection for the \(i\)th observation,
\item \(\mathrm{logit}(p_i)=\log\left(\dfrac{p_i}{1-p_i}\right)\) is the
natural logarithm of the odds of infection,
\item \(\mathbf{x_i}\) is a row vector containing indicator variables for
the variety, nitrogen application timing, and inoculation status of
observation \(i\),
\item \(\mathbf{\beta}\) is a column vector of fixed-effect coefficients, and
\item \(b_{0,j}\sim\mathrm{N}(0,\sigma^2_{b_0})\) and
\(b_{1,j,k}\sim\mathrm{N}(0,\sigma^2_{b_1})\) are random effects for block
and row, respectively.
\end{itemize}

@@

<<models,cache=TRUE>>=
glmm3way.sim <- glmer(cbind(infected, total) ~ variety*n.trt*status +
                        (1|block/row),
                      control = glmerControl(optimizer = 'bobyqa',
                                             optCtrl = list(maxfun = 20000)),
                      family = binomial, data = fert.sim1)
summary(glmm3way.sim)
Anova(glmm3way.sim)

glmm3way <- glmer(cbind(infected, total) ~ variety*n.trt*status +
                    (1|block/row),
                  control = glmerControl(optimizer = 'bobyqa'),
                  family = binomial, data = fert2)
summary(glmm3way)

glmm2way <- glmer(cbind(infected, total) ~ variety*n.trt + variety*status +
                    n.trt*status + (1|block/row),
                    control = glmerControl(optimizer = 'bobyqa'),
                  family = binomial, data = fert2)
summary(glmm2way)
